---
title: "Robust ARIMA Outlier Detection — Bladder Temperature (2h series)"
subtitle: "Bianco/REGARIMA outlier detection per encounter with `robustarima::arima.rob`"
author: ""
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    code-tools: true
    embed-resources: true
    fig-width: 8
    fig-height: 6
execute:
  warning: false
  message: false
  cache: true
params:
  input_csv: "C:/Users/G10073190/Documents/blaas_2h_filled.csv"
  out_dir: "C:/Users/G10073190/Documents"
  run_simulations: false   # case-study injection sweeps (slow!)
---

```{r setup}
#| cache: false
library(dplyr)
library(lubridate)
library(robustarima)
library(ggplot2)

# ---- Parameters ------------------------------------------------------------
INPUT_CSV <- params$input_csv
OUT_DIR   <- params$out_dir

P <- 2; D <- 0; Q <- 2      # ARIMA(p, d, q) for the error model
CRITV   <- 3.0              # outlier critical value (docs: 3 for n < 200)
MIN_OBS <- 50               # skip encounters shorter than this

# arima.rob outlier.type codes -> labels (per package docs)
TYPE_LABEL   <- c("1" = "IO", "2" = "AO", "3" = "LS")
CODE_TO_TYPE <- TYPE_LABEL
TYPE_CODE    <- c(IO = 1, AO = 2, LS = 3)

out_path <- function(...) file.path(OUT_DIR, ...)
```

# Data

```{r load-data}
df <- read.csv(INPUT_CSV, stringsAsFactors = FALSE) %>%
  mutate(
    timestamp = ymd_hms(timestamp),        # ISO timestamps parse directly
    Temp.     = as.numeric(Temp.)
  ) %>%
  filter(!is.na(Temp.)) %>%                # arima.rob disallows missing values
  arrange(encounterId, timestamp)

df %>%
  summarise(
    n_encounters     = n_distinct(encounterId),
    n_rows           = n(),
    rows_per_enc_min = min(table(encounterId)),
    rows_per_enc_med = median(table(encounterId)),
    rows_per_enc_max = max(table(encounterId))
  )
```

# Per-encounter outlier detection

Fits `arima.rob` to one encounter's temperature series and extracts the
detected outliers (additive, level-shift, and innovation). Warnings from the
robustarima package (raised when it judges the flagged contamination rate too
high) are captured per encounter so warned encounters can be reported as
their own category in the summary; the detection output itself is unchanged.

```{r detect}
detect_one <- function(g) {
  eid <- g$encounterId[1]

  if (nrow(g) < MIN_OBS) {
    return(list(
      status = data.frame(encounterId = eid, status = "too_short",
                          n_outliers = NA_integer_,
                          stringsAsFactors = FALSE),
      flags = NULL
    ))
  }

  warn_raised <- FALSE
  fit <- tryCatch(
    withCallingHandlers(
      arima.rob(Temp. ~ 1, data = data.frame(Temp. = g$Temp.),
                p = P, d = D, q = Q,
                innov.outlier = FALSE,
                critv = CRITV, auto.ar = FALSE),
      warning = function(w) {
        warn_raised <<- TRUE
        invokeRestart("muffleWarning")
      }
    ),
    error = function(e) {
      message(sprintf("encounter %s skipped: %s", eid, conditionMessage(e)))
      NULL
    }
  )

  if (is.null(fit)) {
    return(list(
      status = data.frame(encounterId = eid, status = "fit_failed",
                          n_outliers = NA_integer_,
                          stringsAsFactors = FALSE),
      flags = NULL
    ))
  }

  o    <- fit$outliers
  nout <- if (is.null(o)) 0L else o$nout

  flags <- NULL
  if (nout > 0) {
    idx <- o$outlier.index
    flags <- data.frame(
      encounterId = eid,
      timestamp   = g$timestamp[idx],                    # map index -> time
      type        = TYPE_LABEL[as.character(o$outlier.type)],
      impact      = as.numeric(o$outlier.impact),
      t_stat      = as.numeric(o$outlier.t.statistics),
      value       = g$Temp.[idx],
      stringsAsFactors = FALSE
    )
  }

  # Three mutually exclusive categories for processed encounters:
  #   warned  = the robustarima package raised a warning (regardless of flags)
  #   flagged = >= 1 outlier detected, no warning
  #   clean   = 0 outliers detected, no warning
  status <- if (warn_raised) "warned"
            else if (nout > 0) "flagged"
            else "clean"

  list(
    status = data.frame(encounterId = eid, status = status,
                        n_outliers = nout, stringsAsFactors = FALSE),
    flags = flags
  )
}

res <- df %>%
  group_split(encounterId) %>%
  lapply(detect_one)

detect_status <- bind_rows(lapply(res, `[[`, "status"))
outliers      <- bind_rows(lapply(res, `[[`, "flags"))
```

```{r detect-report}
n_enc       <- n_distinct(df$encounterId)
n_processed <- sum(!detect_status$status %in% c("too_short"))
n_failed    <- sum(detect_status$status == "fit_failed")
n_flagged   <- sum(detect_status$status == "flagged")
n_clean     <- sum(detect_status$status == "clean")
n_warned    <- sum(detect_status$status == "warned")

cat(sprintf("encounters total              : %d\n", n_enc))
cat(sprintf("encounters processed          : %d (>= %d obs)\n",
            n_processed, MIN_OBS))
cat(sprintf("  fit failed                  : %d\n", n_failed))
cat(sprintf("  encounters with >=1 anomaly : %d\n", n_flagged))
cat(sprintf("  encounters clean            : %d\n", n_clean))
cat(sprintf("  encounters with warnings    : %d (robustarima warnings)\n",
            n_warned))
cat(sprintf("total outliers flagged        : %d\n", nrow(outliers)))

if (nrow(outliers)) {
  cat("\nbreakdown by type:\n")
  print(table(outliers$type))

  write.csv(outliers, out_path("blaas_outliers.csv"), row.names = FALSE)
  write.csv(detect_status, out_path("blaas_detect_status.csv"),
            row.names = FALSE)
  print(head(outliers, 15))
}
```

# Layer 2 — detection report

Headline = % of encounters with $\ge 1$ detected anomaly; rate = outliers per
observation. The per-outlier detail table is augmented with:

- `se_omega` = SE($\omega$) = |impact| / |t\_stat| (Maronna eq. 8.120/8.122)
- `n_obs` = length of that encounter's series
- `n_interpolated` = interpolated points in that encounter

```{r layer2-report}
# --- Denominators: every PROCESSED encounter (eligible to be flagged) -------
obs_per_enc <- df %>%
  group_by(encounterId) %>%
  summarise(n_obs = n(), .groups = "drop") %>%
  filter(n_obs >= MIN_OBS)

processed_ids <- obs_per_enc$encounterId
n_processed   <- length(processed_ids)
total_obs     <- sum(obs_per_enc$n_obs)

# --- Interpolated-points count per encounter (guard CSV string coercion) ----
interp_per_enc <- df %>%
  mutate(is_interpolated = as.logical(is_interpolated)) %>%
  group_by(encounterId) %>%
  summarise(n_interpolated = sum(is_interpolated, na.rm = TRUE),
            .groups = "drop")

# --- Per-outlier detail table -----------------------------------------------
outliers_aug <- outliers %>%
  mutate(se_omega = abs(impact) / abs(t_stat)) %>%
  left_join(obs_per_enc %>% select(encounterId, n_obs), by = "encounterId") %>%
  left_join(interp_per_enc, by = "encounterId") %>%
  mutate(n_interpolated = coalesce(n_interpolated, 0L)) %>%
  select(encounterId, timestamp, type, impact, t_stat, value,
         se_omega, n_obs, n_interpolated)

write.csv(outliers_aug, out_path("layer2_outliers_detail.csv"),
          row.names = FALSE)

# --- Per-encounter outlier counts (0 for clean encounters) ------------------
per_enc <- obs_per_enc %>%
  left_join(outliers %>% count(encounterId, name = "n_outliers"),
            by = "encounterId") %>%
  left_join(interp_per_enc, by = "encounterId") %>%
  mutate(
    n_outliers       = coalesce(n_outliers, 0L),
    n_interpolated   = coalesce(n_interpolated, 0L),
    has_anomaly      = n_outliers > 0,
    outliers_per_obs = n_outliers / n_obs
  )

# 1. HEADLINE — % of encounters with >= 1 anomaly
n_with   <- sum(per_enc$has_anomaly)
pct_with <- 100 * n_with / n_processed

cat(sprintf("encounters processed        : %d\n", n_processed))
cat(sprintf("encounters with >=1 anomaly : %d (%.1f%%)   <-- HEADLINE\n",
            n_with, pct_with))
cat(sprintf("encounters clean            : %d (%.1f%%)\n",
            n_processed - n_with, 100 - pct_with))

# 2. LENGTH-ADJUSTED RATE — outliers per observation
total_outliers <- nrow(outliers)
rate_overall   <- total_outliers / total_obs
cat(sprintf("\ntotal outliers detected     : %d\n", total_outliers))
cat(sprintf("total observations          : %d\n", total_obs))
cat(sprintf("outliers per observation    : %.4f  (= 1 per %.0f obs)\n",
            rate_overall, 1 / rate_overall))
cat("\nper-encounter outlier rate (outliers / obs):\n")
print(summary(per_enc$outliers_per_obs))

# 3. TYPE BREAKDOWN
type_tbl <- outliers %>%
  count(type, name = "n") %>%
  mutate(pct = round(100 * n / sum(n), 1)) %>%
  arrange(desc(n))
cat("\noutlier type breakdown:\n"); print(type_tbl)

type_enc <- outliers %>%
  distinct(encounterId, type) %>%
  count(type, name = "n_encounters") %>%
  mutate(pct_of_processed = round(100 * n_encounters / n_processed, 1))
cat("\nencounters affected, by type:\n"); print(type_enc)

# 4. SIZE & SIGNIFICANCE + SE, by type
size_tbl <- outliers_aug %>%
  mutate(abs_impact = abs(impact), abs_t = abs(t_stat)) %>%
  group_by(type) %>%
  summarise(
    n          = n(),
    impact_med = round(median(abs_impact), 2),
    impact_iqr = round(IQR(abs_impact), 2),
    impact_max = round(max(abs_impact), 2),
    U_med      = round(median(abs_t), 2),
    U_iqr      = round(IQR(abs_t), 2),
    se_med     = round(median(se_omega), 3),
    se_iqr     = round(IQR(se_omega), 3),
    .groups = "drop"
  )
cat("\nsize (|impact| °C), significance (|U|), and SE by type:\n")
print(size_tbl)

# 5. Per-encounter outlier count (affected only)
cat("\noutliers per encounter (affected encounters only):\n")
print(summary(per_enc$n_outliers[per_enc$has_anomaly]))

write.csv(per_enc,  out_path("layer2_per_encounter.csv"),  row.names = FALSE)
write.csv(type_tbl, out_path("layer2_type_breakdown.csv"), row.names = FALSE)
write.csv(size_tbl, out_path("layer2_size_by_type.csv"),   row.names = FALSE)
```

# Distribution of the outlier t-statistic

Faceted by outlier type (IO/AO/LS). Uses the absolute t-statistic so the
types are compared on magnitude.

```{r plot-tstat}
#| fig-height: 9
#| fig-width: 7
plot_df <- outliers %>%
  filter(!is.na(t_stat)) %>%
  mutate(
    abs_t = abs(t_stat),
    type  = factor(type, levels = c("IO", "AO", "LS"))
  )

type_n <- plot_df %>% count(type)
labeller_n <- setNames(
  sprintf("%s (n = %d)", type_n$type, type_n$n),
  type_n$type
)

p_tstat <- ggplot(plot_df, aes(x = abs_t)) +
  geom_histogram(aes(y = after_stat(density)), bins = 1000,
                 fill = "steelblue", colour = "white", alpha = 0.8) +
  geom_density(colour = "red", linewidth = 0.8) +
  geom_vline(xintercept = 3, linetype = "dashed", colour = "grey40") +
  facet_wrap(~ type, scales = "free", ncol = 1,
             labeller = labeller(type = labeller_n)) +
  labs(
    title = "Distribution of outlier t-statistic by type",
    subtitle = "Dashed line = detection threshold (critv = 3)",
    x = "|t-statistic|", y = "density"
  ) +
  theme_minimal(base_size = 12)

p_tstat
ggsave(out_path("outlier_tstat_by_type.png"), p_tstat,
       width = 7, height = 9, dpi = 150)
```

## Right-tail trimming (per-type IQR cutoff)

Tags each outlier as tail/kept relative to its type's `Q3 + 1.5·IQR` cutoff on
$|t|$, reports the removed tail, and re-plots the kept distribution.

```{r tail-trim}
#| fig-height: 9
#| fig-width: 7
base <- outliers %>%
  filter(!is.na(t_stat)) %>%
  mutate(abs_t = abs(t_stat),
         type  = factor(type, levels = c("IO", "AO", "LS")))

cutoffs <- base %>%
  group_by(type) %>%
  summarise(
    q1     = quantile(abs_t, 0.25),
    q3     = quantile(abs_t, 0.75),
    iqr    = q3 - q1,
    cutoff = q3 + 1.5 * iqr,
    .groups = "drop"
  )
cat("=== Per-type IQR cutoffs on |t_stat| ===\n")
print(cutoffs)

tagged <- base %>%
  left_join(cutoffs %>% select(type, cutoff), by = "type") %>%
  mutate(is_tail = abs_t > cutoff)

kept    <- tagged %>% filter(!is_tail)
removed <- tagged %>% filter(is_tail)

cat(sprintf("\n=== Removed right-tail outliers: %d of %d (%.1f%%) ===\n",
            nrow(removed), nrow(base), 100 * nrow(removed) / nrow(base)))
cat("\nremoved per type:\n")
print(removed %>% count(type, name = "n_removed"))

removed_report <- removed %>%
  select(encounterId, timestamp, type, t_stat, impact, value, cutoff) %>%
  arrange(type, desc(abs(t_stat)))
cat("\nremoved outliers (head):\n")
print(head(removed_report, 20))

write.csv(removed_report, out_path("ttail_removed.csv"), row.names = FALSE)
write.csv(kept %>% select(encounterId, timestamp, type, t_stat, impact, value),
          out_path("ttail_kept.csv"), row.names = FALSE)

kept_n   <- kept %>% count(type)
labels_n <- setNames(sprintf("%s (n = %d)", kept_n$type, kept_n$n), kept_n$type)

p_kept <- ggplot(kept, aes(abs_t)) +
  geom_histogram(bins = 25, fill = "steelblue", colour = "white") +
  facet_wrap(~ type, scales = "free_y", ncol = 1,
             labeller = labeller(type = labels_n)) +
  geom_vline(xintercept = 3, linetype = "dashed", colour = "grey40") +
  labs(
    title = "Outlier t-statistic without right tail",
    subtitle = "Dashed line = detection threshold (critv = 3)",
    x = "|t-statistic|", y = "count"
  ) +
  theme_minimal(base_size = 12)

p_kept
ggsave(out_path("ttail_kept_distribution.png"), p_kept,
       width = 7, height = 9, dpi = 150)
```

# Outlier impact distribution by type

Distribution of $|$impact$|$ (outlier size, in °C) per type, with a log
x-axis to handle the right skew.

```{r plot-impact}
#| fig-height: 9
#| fig-width: 7
plot_df <- outliers %>%
  filter(!is.na(impact)) %>%
  mutate(
    abs_impact = abs(impact),
    type       = factor(type, levels = c("IO", "AO", "LS"))
  ) %>%
  filter(abs_impact > 0)            # log scale can't show zeros

type_n   <- plot_df %>% count(type)
labels_n <- setNames(sprintf("%s (n = %d)", type_n$type, type_n$n),
                     type_n$type)

p_impact <- ggplot(plot_df, aes(abs_impact)) +
  geom_histogram(bins = 20, fill = "darkorange", colour = "white") +
  facet_wrap(~ type, scales = "free_y", ncol = 1,
             labeller = labeller(type = labels_n)) +
  scale_x_log10() +
  labs(
    title = "Outlier impact distribution by type",
    subtitle = "Impact = estimated outlier size (°C), log x-axis",
    x = "|impact| (°C, log scale)", y = "count"
  ) +
  theme_minimal(base_size = 12)

p_impact
ggsave(out_path("impact_by_type.png"), p_impact,
       width = 7, height = 9, dpi = 150)
```

## Significance vs. size (|U| vs |impact|)

Colored by type, because SE($\omega$) — hence the impact→U conversion —
differs by type (LS small SE → steep; AO larger SE → shallower). Dotted lines
are iso-SE references ($U = |\text{impact}| / \text{SE}$).

```{r impact-vs-U}
plot_df <- outliers_aug %>%
  filter(!is.na(impact), !is.na(t_stat)) %>%
  mutate(abs_impact = abs(impact),
         abs_U      = abs(t_stat),
         type       = factor(type, levels = c("AO", "LS")))   # IO off

se_refs <- c(0.1, 0.2, 0.4, 0.8)
xmax <- max(plot_df$abs_impact, na.rm = TRUE)
iso <- do.call(rbind, lapply(se_refs, function(s)
  data.frame(se = factor(s), x = c(0, xmax), y = c(0, xmax / s))))

p_iu <- ggplot(plot_df, aes(abs_impact, abs_U)) +
  geom_line(data = iso, aes(x, y, group = se),
            colour = "grey75", linetype = "dotted", linewidth = 0.4) +
  geom_point(aes(colour = type), alpha = 0.4, size = 1.4) +
  geom_hline(yintercept = 3, linetype = "dashed", colour = "grey40") +
  scale_colour_manual(values = c(AO = "#d62728", LS = "#2ca02c")) +
  coord_cartesian(
    xlim = c(0, quantile(plot_df$abs_impact, 0.99)),
    ylim = c(0, quantile(plot_df$abs_U, 0.99))
  ) +
  labs(
    title = "Outlier significance (|U|) vs. size (|impact|)",
    subtitle = "Dashed line = detection threshold (critv = 3)",
    x = "|impact|  (omega, °C)",
    y = "|U|  (detection statistic)",
    colour = "type"
  ) +
  theme_minimal(base_size = 12)

p_iu
ggsave(out_path("impact_vs_U.png"), p_iu, width = 8, height = 6, dpi = 150)
```

# Per-encounter series plot

Plots one encounter's temperature series, marks detected outliers by type,
and annotates each with its t-statistic.

```{r plot-encounter-fn}
TYPE_COL   <- c(IO = "#1f77b4", AO = "#d62728", LS = "#2ca02c")
TYPE_SHAPE <- c(IO = 3, AO = 2, LS = 5)

plot_encounter_outliers <- function(enc_id, df, outliers, save_dir = NULL) {
  series <- df %>%
    filter(encounterId == enc_id) %>%
    arrange(timestamp)
  if (nrow(series) == 0) stop(sprintf("no data for encounter %s", enc_id))

  flags <- outliers %>%
    filter(encounterId == enc_id) %>%
    mutate(type = factor(type, levels = names(TYPE_COL)))

  p <- ggplot(series, aes(timestamp, Temp.)) +
    geom_line(colour = "grey50", linewidth = 0.5) +
    geom_point(colour = "grey30", size = 1)

  if (nrow(flags) > 0) {
    p <- p +
      geom_point(data = flags,
                 aes(timestamp, value, colour = type, shape = type),
                 size = 4, stroke = 1.1) +
      geom_text(data = flags,
                aes(timestamp, value,
                    label = sprintf("t=%.2f", t_stat), colour = type),
                vjust = 2.2, size = 3, show.legend = FALSE) +
      scale_colour_manual(values = TYPE_COL, drop = FALSE, name = "type") +
      scale_shape_manual(values = TYPE_SHAPE, drop = FALSE, name = "type")
  }

  p <- p +
    labs(
      title = sprintf("Encounter %s — temperature with detected outliers", enc_id),
      x = "time", y = "Temperatuur (°C)"
    ) +
    theme_minimal(base_size = 12)

  if (!is.null(save_dir)) {
    f <- file.path(save_dir, sprintf("encounter_%s_outliers.png", enc_id))
    ggsave(f, p, width = 11, height = 5, dpi = 150)
  } else {
    print(p)
  }
  invisible(p)
}
```

```{r plot-encounter-example}
#| fig-width: 11
#| fig-height: 5
plot_encounter_outliers(enc_id = 30775, df = df, outliers = outliers)

# Or loop over every encounter that had an outlier, saving each:
# for (eid in unique(outliers$encounterId))
#   plot_encounter_outliers(eid, df, outliers,
#                           save_dir = out_path("outlier_plots"))
```

# Case studies — AO injection on 10 clean series

Both studies inject **additive outliers (AO) only**, with random sign per
injected point, into the **same ten clean encounters**, and run the **exact
production detector**: fixed ARIMA(`r P`, `r D`, `r Q`),
`innov.outlier = FALSE`, `critv = `r CRITV``, `auto.ar = FALSE`.

- **Study 1 — reliability.** At a fixed 5% contamination, AOs of magnitude
  `{0.3, 0.5, 1.0, 1.5, 2.0}` °C are injected and swept. Question: how
  reliably does the detector recover injected outliers across magnitudes,
  and how many false positives does it produce along the way? Within a
  series the injection locations are held fixed across magnitudes, so the
  magnitude effect is isolated.
- **Study 2 — swamping.** At a fixed magnitude (set below), contamination is
  swept over `{5, 10, 15, 25}`%. Question: do false positives rise as the
  series gets more contaminated? Locations are drawn fresh at every
  contamination level.

Every injected point is classified as **recovered** (AO detected at the
location), **misspecified** (an outlier of another type detected there), or
**missed** (nothing detected there). Any detection away from all injected
locations counts as a **false positive**.

::: {.callout-note}
These sweeps re-fit `arima.rob` many times. They run only when
`params$run_simulations` is `TRUE` — render with
`quarto render blaas_outlier_analysis.qmd -P run_simulations:true`.
:::

```{r case-setup}
#| eval: !expr params$run_simulations
# >>> FILL IN: ten clean encounter IDs for the case studies <<<
CASE_IDS <- c()          # e.g. c(5729, 6011, ...)

# Clean = processed, no flags, no robustarima warning
clean_ids <- detect_status %>%
  filter(status == "clean") %>%
  pull(encounterId)

if (length(CASE_IDS) == 0) {
  CASE_IDS <- head(clean_ids, 10)
  cat("CASE_IDS not set — defaulting to the first 10 clean encounters:\n")
}
stopifnot(all(CASE_IDS %in% detect_status$encounterId))
not_clean <- setdiff(CASE_IDS, clean_ids)
if (length(not_clean) > 0)
  warning("these CASE_IDS are not in the clean pool: ",
          paste(not_clean, collapse = ", "))
print(CASE_IDS)

CS_TOL    <- 1                              # location tolerance for a "hit"
CS_MAGS   <- c(0.3, 0.5, 1.0, 1.5, 2.0)     # study 1: magnitude sweep (°C)
CS_CONTAM <- c(0.05, 0.10, 0.15, 0.25)      # study 2: contamination sweep

# The EXACT production detector (same spec as detect_one above)
fit_detect <- function(y) {
  withCallingHandlers(
    arima.rob(Temp. ~ 1, data = data.frame(Temp. = y),
              p = P, d = D, q = Q,
              innov.outlier = FALSE, critv = CRITV, auto.ar = FALSE),
    warning = function(w) invokeRestart("muffleWarning")
  )
}

inject_AO_multi <- function(y, locs, signs, mag) {
  y[locs] <- y[locs] + signs * mag
  y
}

# Classify every injected location and collect off-location false positives.
classify_fit_result <- function(fit, locs, tol = CS_TOL) {
  o <- fit$outliers
  det_idx  <- integer(0); det_type <- character(0)
  if (!is.null(o) && o$nout > 0) {
    det_idx  <- o$outlier.index
    det_type <- unname(CODE_TO_TYPE[as.character(o$outlier.type)])
  }
  detected <- vapply(locs, function(L) {
    hit <- which(abs(det_idx - L) <= tol)
    if (!length(hit)) "none"
    else if (any(det_type[hit] == "AO")) "AO"
    else det_type[hit[1]]                     # misspecified type
  }, character(1))
  is_fp <- if (length(det_idx))
    vapply(det_idx, function(i) all(abs(i - locs) > tol), logical(1))
  else logical(0)
  list(detected = detected,
       fp_type  = det_type[is_fp],
       n_fp     = sum(is_fp))
}

# One run = one series x one contamination fraction x one magnitude.
# The seed determines the injection locations:
#   study 1 uses ONE seed per series -> same locations across magnitudes
#   study 2 uses a seed per (series, level) -> fresh locations per level
run_case <- function(eid, contam_frac, mag, seed_offset = 0) {
  set.seed(eid + seed_offset)
  y0 <- df$Temp.[df$encounterId == eid]
  n  <- length(y0)
  n_inject <- max(1, round(contam_frac * n))
  locs     <- sort(sample(seq(3, n - 2), n_inject))
  signs    <- sample(c(1, -1), n_inject, replace = TRUE)

  y_inj <- inject_AO_multi(y0, locs, signs, mag)
  fit   <- tryCatch(fit_detect(y_inj), error = function(e) NULL)

  if (is.null(fit)) {
    return(list(
      locations = data.frame(encounterId = eid, n_obs = n,
                             contam_target = contam_frac,
                             contam_actual = n_inject / n,
                             magnitude = mag, location = locs,
                             injected = "AO", detected = NA_character_),
      summary = data.frame(encounterId = eid, n_obs = n,
                           contam_target = contam_frac,
                           contam_actual = n_inject / n,
                           magnitude = mag, n_injected = n_inject,
                           n_recovered = NA, n_misspec = NA, n_missed = NA,
                           n_fp = NA, fit_ok = FALSE)
    ))
  }

  cl <- classify_fit_result(fit, locs)
  list(
    locations = data.frame(encounterId = eid, n_obs = n,
                           contam_target = contam_frac,
                           contam_actual = n_inject / n,
                           magnitude = mag, location = locs,
                           injected = "AO", detected = cl$detected),
    fp = if (cl$n_fp > 0)
      data.frame(encounterId = eid, n_obs = n,
                 contam_target = contam_frac, magnitude = mag,
                 fp_type = cl$fp_type) else NULL,
    summary = data.frame(
      encounterId = eid, n_obs = n,
      contam_target = contam_frac, contam_actual = n_inject / n,
      magnitude = mag, n_injected = n_inject,
      n_recovered = sum(cl$detected == "AO"),
      n_misspec   = sum(!cl$detected %in% c("AO", "none")),
      n_missed    = sum(cl$detected == "none"),
      n_fp        = cl$n_fp, fit_ok = TRUE)
  )
}

collect_runs <- function(runs) {
  list(
    locations = bind_rows(lapply(runs, `[[`, "locations")),
    fp        = bind_rows(lapply(runs, `[[`, "fp")),
    summary   = bind_rows(lapply(runs, `[[`, "summary"))
  )
}
```

## Study 1 — reliability: magnitude sweep at 5% contamination

```{r case-study1}
#| eval: !expr params$run_simulations
STUDY1_SEED <- 1   # one seed per series -> locations fixed across magnitudes

grid1 <- expand.grid(encounterId = CASE_IDS, mag = CS_MAGS) %>%
  mutate(contam = 0.05)

study1 <- collect_runs(
  Map(function(eid, mg) run_case(eid, 0.05, mg, STUDY1_SEED),
      grid1$encounterId, grid1$mag)
)

cat("=== Study 1: recovery & false positives by magnitude ===\n")
study1$summary %>%
  filter(fit_ok) %>%
  group_by(magnitude) %>%
  summarise(
    n_series      = n_distinct(encounterId),
    recovery_rate = sum(n_recovered) / sum(n_injected),
    miss_rate     = sum(n_missed)    / sum(n_injected),
    misspec_rate  = sum(n_misspec)   / sum(n_injected),
    fp_per_series = mean(n_fp),
    fp_rate       = sum(n_fp) / sum(n_obs - n_injected),
    .groups = "drop"
  ) %>%
  print(digits = 3)

cat("\n=== Study 1: confusion matrix (injected AO vs detected type) ===\n")
print(with(study1$locations, table(injected, detected, useNA = "ifany")))

cat("\n=== Study 1: confusion by magnitude (row-normalized) ===\n")
print(round(prop.table(
  with(study1$locations, table(magnitude, detected)), margin = 1), 3))

if (nrow(study1$fp) > 0) {
  cat("\n=== Study 1: false-positive flags by type ===\n")
  print(with(study1$fp, table(magnitude, fp_type)))
}
```

## Study 2 — swamping: contamination sweep at fixed magnitude

```{r case-study2}
#| eval: !expr params$run_simulations
# >>> CHOOSE the injected AO magnitude (°C) here <<<
CS_MAG_FIXED <- 0.8

# Same ten series as study 1; fresh locations per contamination level.
study2_seed <- function(contam) 2000 + round(1000 * contam)

grid2 <- expand.grid(encounterId = CASE_IDS, contam = CS_CONTAM)

study2 <- collect_runs(
  Map(function(eid, cf) run_case(eid, cf, CS_MAG_FIXED, study2_seed(cf)),
      grid2$encounterId, grid2$contam)
)

cat(sprintf("=== Study 2: false positives by contamination (magnitude %.2f °C) ===\n",
            CS_MAG_FIXED))
n_failed_s2 <- sum(!study2$summary$fit_ok)
if (n_failed_s2 > 0)
  cat(sprintf("(fits failed: %d — excluded from the rates below)\n", n_failed_s2))
study2$summary %>%
  filter(fit_ok) %>%
  group_by(contam_target) %>%
  summarise(
    n_series      = n_distinct(encounterId),
    recovery_rate = sum(n_recovered) / sum(n_injected),
    fp_per_series = mean(n_fp),
    fp_rate       = sum(n_fp) / sum(n_obs - n_injected),
    .groups = "drop"
  ) %>%
  print(digits = 3)

cat("\n=== Study 2: confusion matrix (injected AO vs detected type) ===\n")
print(with(study2$locations, table(injected, detected, useNA = "ifany")))

cat("\n=== Study 2: confusion by contamination level (row-normalized) ===\n")
print(round(prop.table(
  with(study2$locations, table(contam_target, detected)), margin = 1), 3))

if (nrow(study2$fp) > 0) {
  cat("\n=== Study 2: false-positive flags by type ===\n")
  print(with(study2$fp, table(contam_target, fp_type)))
}
```

## Two-panel rate plot (both studies)

Left panel: rates vs injected magnitude (Study 1, 5% contamination). Right
panel: rates vs contamination level (Study 2, fixed magnitude). Lines:
recovered / misspecified / missed as shares of injected points, and false
positives per non-injected observation.

```{r case-plot}
#| eval: !expr params$run_simulations
#| fig-width: 10
#| fig-height: 4.5
rates_from <- function(smry, xvar, xlab_fmt) {
  smry %>%
    filter(fit_ok) %>%
    group_by(x = .data[[xvar]]) %>%
    summarise(
      recovered      = sum(n_recovered) / sum(n_injected),
      misspecified   = sum(n_misspec)   / sum(n_injected),
      missed         = sum(n_missed)    / sum(n_injected),
      false_positive = sum(n_fp) / sum(n_obs - n_injected),
      .groups = "drop"
    ) %>%
    tidyr::pivot_longer(-x, names_to = "outcome", values_to = "rate") %>%
    mutate(x_lab = xlab_fmt(x))
}

pd <- bind_rows(
  rates_from(study1$summary, "magnitude",
             function(x) sprintf("%.1f", x)) %>%
    mutate(panel = "Outlier magnitude (°C)"),
  rates_from(study2$summary, "contam_target",
             function(x) sprintf("%.0f%%", 100 * x)) %>%
    mutate(panel = "Contamination rate")
) %>%
  mutate(
    panel   = factor(panel, levels = c("Outlier magnitude (°C)",
                                       "Contamination rate")),
    outcome = factor(outcome, levels = c("recovered", "misspecified",
                                         "missed", "false_positive")),
    x_lab   = factor(x_lab, levels = unique(x_lab))
  )

OUTCOME_COL <- c(recovered      = "#2ca02c",
                 misspecified   = "#9467bd",
                 missed         = "#1f77b4",
                 false_positive = "#d62728")
OUTCOME_LTY <- c(recovered      = "solid",
                 misspecified   = "dashed",
                 missed         = "dotted",
                 false_positive = "dashed")

p_case <- ggplot(pd, aes(x_lab, rate, colour = outcome,
                         linetype = outcome, group = outcome)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.8) +
  facet_wrap(~ panel, scales = "free_x") +
  scale_colour_manual(values = OUTCOME_COL, name = NULL) +
  scale_linetype_manual(values = OUTCOME_LTY, name = NULL) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "rate",
       title = "AO recovery, misspecification, and false positives — production detector",
       subtitle = sprintf(
         "Left: 5%% contamination, magnitude swept | Right: magnitude %.2f °C, contamination swept | 10 clean series",
         CS_MAG_FIXED)) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())

p_case
ggsave(out_path("case_rates_two_panel.png"), p_case,
       width = 10, height = 4.5, dpi = 150)
```

## Series plots — one series per condition

The first case series (`CASE_IDS[1]`) is shown at every condition: 5 plots
for the Study 1 magnitudes, 4 for the Study 2 contamination levels (9 total).
The seeds match the studies exactly, so the plots show the very injections
that produced the numbers above. Uncomment the loop at the bottom to save
plots for all ten series.

```{r case-series-fn}
#| eval: !expr params$run_simulations
plot_case_series <- function(eid, contam_frac, mag, seed_offset) {
  set.seed(eid + seed_offset)
  enc <- df %>% filter(encounterId == eid) %>% arrange(timestamp)
  y0  <- enc$Temp.; ts <- enc$timestamp; n <- length(y0)
  n_inject <- max(1, round(contam_frac * n))
  locs     <- sort(sample(seq(3, n - 2), n_inject))
  signs    <- sample(c(1, -1), n_inject, replace = TRUE)
  y_inj    <- inject_AO_multi(y0, locs, signs, mag)

  fit <- tryCatch(fit_detect(y_inj), error = function(e) NULL)
  if (is.null(fit)) {
    cat(sprintf("eid %s | contam %.0f%% | mag %.1f: fit failed\n",
                eid, 100 * contam_frac, mag))
    return(invisible(NULL))
  }
  cl <- classify_fit_result(fit, locs)

  o <- fit$outliers
  det_idx <- if (!is.null(o) && o$nout > 0) o$outlier.index else integer(0)
  fp_idx  <- det_idx[vapply(det_idx, function(i) all(abs(i - locs) > CS_TOL),
                            logical(1))]

  series_df <- data.frame(timestamp = ts, temp = y_inj)
  inj_df    <- data.frame(timestamp = ts[locs], temp = y_inj[locs],
                          outcome = ifelse(cl$detected == "AO", "recovered",
                                    ifelse(cl$detected == "none", "missed",
                                           "misspecified")))
  fp_df <- if (length(fp_idx)) data.frame(timestamp = ts[fp_idx],
                                          temp = y_inj[fp_idx]) else NULL

  p <- ggplot(series_df, aes(timestamp, temp)) +
    geom_line(colour = "grey55", linewidth = 0.4) +
    geom_point(colour = "grey35", size = 0.8) +
    geom_point(data = inj_df, aes(colour = outcome),
               shape = 1, size = 3.2, stroke = 1.2) +
    scale_colour_manual(values = c(recovered = "#2ca02c",
                                   misspecified = "#9467bd",
                                   missed = "#d62728"),
                        name = "injected AO", drop = FALSE) +
    labs(title = sprintf(
           "Encounter %s — AO injection, magnitude %.1f °C, %.0f%% contamination",
           eid, mag, 100 * contam_frac),
         subtitle = sprintf("n = %d | %d injected | × = false-positive flag",
                            n, n_inject),
         x = "time", y = "Temp (°C)") +
    theme_minimal(base_size = 11)
  if (!is.null(fp_df))
    p <- p + geom_point(data = fp_df, shape = 4, size = 3,
                        colour = "black", stroke = 1)
  print(p)
  ggsave(out_path(sprintf("case_series_%s_mag%.1f_c%.0f.png",
                          eid, mag, 100 * contam_frac)),
         p, width = 10, height = 4.5, dpi = 150)
  invisible(p)
}
```

```{r case-series-study1}
#| eval: !expr params$run_simulations
#| fig-width: 10
#| fig-height: 4.5
# Study 1 conditions: first series at each magnitude (5% contamination)
for (m in CS_MAGS)
  plot_case_series(CASE_IDS[1], 0.05, m, STUDY1_SEED)
```

```{r case-series-study2}
#| eval: !expr params$run_simulations
#| fig-width: 10
#| fig-height: 4.5
# Study 2 conditions: first series at each contamination level
for (cf in CS_CONTAM)
  plot_case_series(CASE_IDS[1], cf, CS_MAG_FIXED, study2_seed(cf))

# All ten series at every condition (saves 90 PNGs):
# for (eid in CASE_IDS) {
#   for (m in CS_MAGS)   plot_case_series(eid, 0.05, m, STUDY1_SEED)
#   for (cf in CS_CONTAM) plot_case_series(eid, cf, CS_MAG_FIXED, study2_seed(cf))
# }
```

# Session info

```{r session-info}
#| cache: false
sessionInfo()
```
